Testing the Relationship between Phenomenological Control related to Illusion Sensitivity

### 2024-06-09 paper/R0/manuscript.qmd
submitted

To encourage the empirical exploration of our ability and tendency to alter our phenomenological experience and further accelerate investigations away from the hypnotic context, @lush2021 adapted the Sussex-Waterloo Scale of Hypnotisability [SWASH, @lush2018] by removing all its references to hypnosis, to measure trait phenomenological control. This newly developed phenomenological control scale (PCS) consists of 10 imaginative suggestions followed by subjective ratings for each suggestion on a 6-point Likert scale and has been demonstrated to be compatible with online experiments [@lush2022].

Interestingly, @lush2022 did test for a relationship between PC and illusion sensitivity using the Müller-Lyer illusion (in which the arrangement of the arrowheads flanking two lines makes them appear as having different lengths), and reported evidence in favour of an absence of correlation between the two measures. This finding was interpreted as indicative of the cognitive impenetrability of illusions, implying that the effect is driven by low-level processes and therefore not influenced by top-down mechanisms such as PC. The goal of this study is thus to replicate the results from @lush2022 pointing to an absence of a relationship between phenomenological control and illusion sensitivity, by generalising them to a different illusion task that encompasses other illusion types (see @tbl-table1).

>>>>

To encourage the empirical exploration of our ability and tendency to alter our phenomenological experience and further accelerate investigations away from the hypnotic context, @lush2021 adapted the Sussex-Waterloo Scale of Hypnotisability [SWASH, @lush2018] by removing all its references to hypnosis, to measure trait phenomenological control. This newly developed phenomenological control scale (PCS) consists of 10 imaginative suggestions followed by subjective ratings for each suggestion on a 6-point Likert scale and has demonstrated validity in online experiments [@lush2022].

Interestingly, @lush2022 did test for a relationship between PC and illusion sensitivity using the Müller-Lyer illusion (in which the arrangement of the arrowheads flanking two lines makes them appear as having different lengths), and reported evidence in favour of an absence of correlation between the two measures. This finding was interpreted as indicative of the cognitive impenetrability of illusions, implying that the effect is driven by low-level processes and therefore not influenced by top-down mechanisms such as PC. The goal of this study is thus to replicate the results from @lush2022 pointing to an absence of a relationship between phenomenological control and illusion sensitivity, by generalising them to a different illusion paradigm that encompasses other illusion types (see @tbl-table1).
